### 2025-05-16 manuscript/manuscript.qmd
second commit, all main analysis done

---
title: ""
bibliography: "misc/ref.bib"
csl: "misc/vancouver.csl"
format: 
  docx: 
    reference-doc: "ref.docx"
---

**Title**: Different Cardiac Autonomic Trajectories in Response to Exercise-Induced Stress

**Authors**: Matías Castillo-Aguilar,^1^ ..., Cristian Núñez-Espinosa,^1,2^ \*

::: {custom-style="Afiliation"}
^1^ Department/Affiliation, University/Institution, City, State/Province, Country

^2^ Department/Affiliation, University/Institution, City, State/Province, Country
:::

### \*Correspondence

Cristian Núñez-Espinosa, Escuela de Medicina, Universidad de Magallanes (UMAG), Punta Arenas, Chile. e-mail: [cristian.nunez\@umag.cl](mailto:cristian.nunez@umag.cl). Address: [Full Address]. Phone: [Phone Number].

{{< pagebreak >}}

## Abstract

**Objective**: Briefly state the primary aim(s) of your research investigation.

**Material and methods**: Briefly describe the overall research design, participants/samples, procedures, and data analysis methods.

**Results**: Summarize the key findings of your study, including relevant data and statistical results.

**Conclusion**: State the main conclusions drawn from your research and their significance.

**Keywords**: List 3-5 keywords relevant to your research topic.

{{< pagebreak >}}

# Introduction

<!--- 
OPENING HOOK SUGGESTION: 
Start with 2-3 sentences that link a clinically relevant syndrome/condition to an unresolved mechanistic question. 
- Define the condition using 1-2 seminal citations. 
- Highlight its epidemiological or public health burden. 
- Introduce the core system/mechanism under investigation as the study's focus.
--->
[opening paragraph here...]

<!--- 
## 1. **Mechanistic Context & Knowledge Gaps**

### 1.1. Foundational Understanding

LAYER 1 SUGGESTION: 
Summarize established foundational knowledge about the system/mechanism. 
- Cite 2-3 key papers that define current consensus. 
- Mention widely used metrics/tools in the field (e.g., "[METRIC] measures [PHENOMENON]").
--->
[established knowledge summary...]

<!--- 
### 1.2. Emerging Contradictions & Unanswered Questions

LAYER 2 SUGGESTION: 
Highlight emerging contradictions or unanswered questions. Use contrastive language:
- "While [X] is well-documented, [Y] remains controversial..."
- "Recent studies suggest [Z], but fail to address [GAP]..."
--->
[gap identification...]

<!--- 
### 1.3. Methodological Limitations

LAYER 3 SUGGESTION: 
Critique methodological limitations of current approaches. 
- Explain why traditional tools/metrics are insufficient (e.g., "Static measures overlook [DYNAMIC PROCESS]"). 
- Introduce the need for innovative methods (e.g., "[NEW METHOD] better captures [COMPLEXITY]").
--->
[methodological critique...]

<!--- 
## 2. **Methodological Rationale & Hypothesis**

### 2.1. Justification of Research Paradigm

TOOL 1 SUGGESTION: 
Justify your ecologically/clinically relevant paradigm. 
- Compare it to lab-based alternatives (e.g., "[REAL-WORLD ASSAY] mimics [CONDITION] better than [TRADITIONAL ASSAY]"). 
- Cite validation studies if available.
--->
[paradigm justification...]

<!--- 
### 2.2. Innovative Analytical/Metric Approach

TOOL 2 SUGGESTION: 
Explain your novel analytical/metric approach. 
- Highlight how it addresses Layer 3 limitations (e.g., "[COMPOSITE INDEX] integrates [MULTI-SYSTEM SIGNALS]"). 
- Cite technical validation papers.
--->
[method explanation...]

<!--- 
### 2.3. Hypothesis Formulation

HYPOTHESIS SUGGESTION: 
Explicitly state how your approach resolves the gap. Use the formula:
"By combining [TOOL 1] with [TOOL 2], this study tests the hypothesis that [CONDITION] disrupts [MECHANISM] through [PATHWAY]."
--->
[hypothesis statement...]

<!--- 
## 3. **Anticipated Impact & Translation**

### 3.1. Mechanistic Impact

MECHANISTIC IMPACT SUGGESTION: 
Explain how findings could refine fundamental understanding. Example:
"Clarifying [X] may disentangle [Y] from [Z] in [SYSTEM], resolving long-standing debates about [CONTROVERSY]."
--->
[mechanistic impact...]

<!--- 
### 3.2. Clinical Translation

CLINICAL TRANSLATION SUGGESTION: 
Link mechanisms to actionable insights. Example:
"Identifying [BIOMARKER/PATHWAY] could guide [INTERVENTION] for [SUBGROUP], addressing [UNMET CLINICAL NEED]."
--->
[clinical translation...]

<!--- 
### 3.3. Paradigm Shift in Methodology

PARADIGM SHIFT SUGGESTION: 
Highlight broader methodological implications. Example:
"This [METHODOLOGY] bridges [DISCIPLINE 1] and [DISCIPLINE 2], enabling future studies of [COMPLEX PHENOMENA]."
--->
[paradigm shift...]

<!--- 
## 4. **Additional Considerations**

### 4.1. Rhetorical Devices & Integration

RHETORICAL DEVICES SUGGESTION: 
- Use citation clusters thematically (e.g., group mechanistic studies separately from clinical papers). 
- Employ contrastive syntax to emphasize gaps ("Although...", "However..."). 
- Connect methods to hypotheses with bridging language ("To resolve this, we...").
--->
[narrative integration and rhetorical devices...]

<!--- 
Final Checklist for Blueprint Consistency

FINAL CHECKLIST: 
- [ ] Hook ties condition to mechanism in ≤3 sentences 
- [ ] Gap narrative progresses: Consensus → Contradictions → Method limitations 
- [ ] Hypothesis explicitly links tools to the gap 
- [ ] Impact addresses both basic science and translation
--->

# Material and Methods

## Study Design

### Overall Study Framework  
<!---  
Describe the general design of your study (e.g., experimental, observational, clinical trial, cohort study, cross-sectional study, case-control study).  
Provide context for why this study type is most appropriate for addressing your research question.  
--->
[Your overall design description...]

### Structural Specifics and Justification  
<!---  
Detail the specific structure of the study (e.g., randomized, placebo-controlled, blind, longitudinal, cross-sectional).  
Explain the rationale behind choosing this structure, emphasizing its advantages over alternatives.  
Include any necessary visual aids such as a flowchart to illustrate the study design (optional but recommended).  
--->
[Your detailed design and justification...]

## Participants

### Recruitment and Sampling Strategy  
<!---  
Describe the methods used for recruitment or sample acquisition (e.g., clinics, community outreach, database screening).  
Outline the process for selecting participants to ensure a representative and unbiased sample.  
--->
[Your recruitment details...]

### Inclusion and Exclusion Criteria  
<!---  
Specify clear inclusion and exclusion criteria for participants or samples to ensure the study’s internal validity.  
Discuss any stratification or subgroup definitions if applicable.  
--->
[Your inclusion/exclusion criteria...]

### Sample Size and Demographics  
<!---  
Provide the sample size and justification for its adequacy (e.g., through power analysis).  
Describe key demographic characteristics (age, sex, ethnicity, etc.) and any variability considerations.  
--->
[Your sample size rationale and demographic details...]

### Ethical Considerations  
<!---  
Document the ethical aspects of the study, including:  
- Institutional review board (IRB) or ethics committee approval.  
- Informed consent procedures.  
- Compliance with animal welfare practices (if applicable).  
--->
[Your ethical procedures...]

## Procedures

### Experimental Protocols and Interventions  
<!---  
Provide a comprehensive description of all experimental protocols, treatments, interventions, and control conditions.  
Detail key variables such as dosage, treatment duration, frequency, and administration method where relevant.  
--->
[Your experimental procedures...]

### Standardization and Instrumentation  
<!---  
Explain the instruments or equipment used for measurements, including calibration and standardization procedures to ensure reproducibility.  
Include details regarding the setup, maintenance, and quality assurance of these instruments.  
--->
[Your instrumentation details...]

## Assessments

### Assessment Strategy  
<!---  
Outline the assessments conducted as part of the study.  
Explain how each assessment aligns with the study objectives and outcomes of interest.  
--->
[Your overall assessment strategy...]

### Assessment Protocols

#### Assessment 1

<!---  
Describe the specific variable(s) being measured in the first assessment.  
Detail the instrument(s) or techniques used, along with any calibration or validation steps.  
Include quality control measures to ensure data reliability.  
--->
[Detailed description of Assessment 1.]

#### Assessment 2

<!---  
Elaborate on the second assessment using the same structured approach:  
- Variable(s) measured  
- Instrumentation and methodology  
- Quality control and validation procedures  
--->
[Detailed description of Assessment 2.]

#### Assessment 3

<!---  
Provide a similar detailed protocol for a third assessment if applicable.  
--->
[Detailed description of Assessment 3.]

## Statistical Analysis

<!---  
### 5.1. Data Analysis Methodology

Describe the overall statistical methods used to analyze the data.  
List the specific statistical tests applied (e.g., Student's t-test, ANOVA, linear regression, etc.).  
Outline how data distributions were assessed and any transformations applied.  
--->
[Your detailed data analysis methodology...]

<!---  
### 5.2. Software and Significance Criteria

Specify the statistical software and programming languages used (e.g., R, SPSS, SAS).  
Indicate the level of statistical significance (e.g., p < 0.05) and any adjustments for multiple comparisons if applicable.  
--->
[Your software and significance details...]

<!---  
### 5.3. Handling Confounders and Robustness Checks

Discuss the methods used to control for potential confounding variables.  
Include any sensitivity analyses or data robustness checks implemented to validate the results.  
--->
[Your methods to control confounders...]

<!---  
### 5.4. Data Quality Assurance

Outline the quality control procedures and verification methods for ensuring data integrity.  
This may include repeated measures, inter-rater reliability assessments, or external validation steps.  
--->
[Your quality assurance protocols...]

# Results

-   Present the results in a clear, concise, and logical manner.
-   Use tables and figures to effectively visualize the data. Ensure tables and figures are numbered and have descriptive captions.
-   Report relevant statistical results, including p-values, effect sizes, and confidence intervals.
-   Present the main findings first, followed by secondary findings.
-   Avoid repeating data in the text that is already presented in tables or figures, but highlight the most important trends and findings.

# Discussion

<!--- 
OPENING HOOK SUGGESTION: 
Begin with 2-3 sentences that succinctly recapitulate the study’s principal findings. 
- Frame the findings within the context of the initial research question. 
- Emphasize the relevance of these results to current debates or clinical challenges.
--->
[opening paragraph here...]

<!--- 
## 1. **Interpretation & Integration of Findings**

### 1.1. Summary of Key Results

SUMMARY SUGGESTION: 
Recap the major outcomes of the study in a succinct manner. 
- Highlight results that directly address the hypothesis.
- Mention any quantitative or qualitative shifts observed.
--->
[summary of findings...]

<!--- 
### 1.2. Comparative Analysis with Existing Literature

COMPARISON SUGGESTION: 
Contrast your findings with established studies. 
- Identify similarities and divergences relative to prior reports.
- Use precise citations to underline both agreements and differences.
--->
[literature comparison...]

<!--- 
### 1.3. Elucidation of Underlying Mechanisms

MECHANISTIC INTERPRETATION SUGGESTION: 
Delve into the mechanistic implications of your findings. 
- Explain how the data inform or challenge prevailing theories.
- Highlight any novel mechanistic insights that emerge.
--->
[mechanistic interpretation...]

<!--- 
## 2. **Contextualizing Findings Within the Field**

### 2.1. Alignment with Existing Models

MODEL INTEGRATION SUGGESTION: 
Discuss how your results fit within current conceptual frameworks. 
- Compare your approach and outcomes with traditional and emerging models.
- Provide scholarly context through critical references.
--->
[analysis of model integration...]

<!--- 
### 2.2. Addressing Discrepancies and Controversies

CONTRAST & CONTROVERSY SUGGESTION: 
Identify and elaborate on any inconsistencies between your data and previous studies. 
- Use contrastive language to articulate divergent findings.
- Discuss possible reasons (e.g., methodological differences or biological variability).
--->
[discussion on discrepancies...]

<!--- 
## 3. **Implications for Future Research**

### 3.1. Methodological Refinements and Innovations

METHODOLOGICAL INSIGHTS SUGGESTION: 
Critically examine the techniques used in your study. 
- Discuss the strengths and limitations of your chosen methods.
- Suggest potential improvements or alternative approaches for future studies.
--->
[methodological insights...]

<!--- 
### 3.2. Proposed Future Directions

FUTURE OUTLOOK SUGGESTION: 
Outline new research questions or experiments that arise from your findings. 
- Highlight gaps that need further investigation.
- Propose additional studies to validate and extend your current work.
--->
[proposed future research directions...]

<!--- 
## 4. **Limitations and Alternative Interpretations**

### 4.1. Technical and Experimental Constraints

LIMITATIONS SUGGESTION: 
Provide a candid discussion of the study’s limitations. 
- Address technical and methodological issues that may impact interpretation.
- Reflect on sample size, assay specificity, or other constraints.
--->
[limitations discussion...]

<!--- 
### 4.2. Impact on Data Interpretation and Alternate Models

ALTERNATIVE INTERPRETATIONS SUGGESTION: 
Discuss how these limitations might shape alternative views or interpretations of the data. 
- Consider other variables or confounders that could explain your findings.
- Suggest ways to mitigate these issues in follow-up studies.
--->
[analysis of alternative interpretations...]

<!--- 
## 5. **Concluding Synthesis and Translational Impact**

### 5.1. Synthesis of Discussion Points

SYNTHESIS SUGGESTION: 
Recap the integration of your findings with the broader literature. 
- Summarize key insights and the advancement in mechanistic understanding.
--->
[synthesis of discussion points...]

<!--- 
### 5.2. Clinical and Translational Relevance

CLINICAL IMPLICATIONS SUGGESTION: 
Highlight the practical implications of your findings for patient care or intervention strategies.
- Discuss how these insights might influence clinical decision-making or therapeutic development.
--->
[clinical implications...]

<!--- 
### 5.3. Broader Impact and Future Perspectives

BROADER IMPLICATIONS SUGGESTION: 
Emphasize the potential paradigm shifts your study may foster.
- Outline how your findings could influence related fields or future multidisciplinary approaches.
--->
[broader impact discussion...]

<!--- 
## 6. **Additional Considerations & Final Checklist**

### 6.1. Integration of Rhetorical Strategies

RHETORICAL DEVICES SUGGESTION: 
Detail the use of linking language and structured argumentation to weave together contrasting points.
- Emphasize clarity and logical progression throughout the discussion.
--->
[commentary on rhetorical strategies...]

<!--- 
Final Checklist for Discussion Section Integrity

FINAL CHECKLIST SUGGESTION: 
- [ ] The discussion begins with a concise recap of key findings.
- [ ] Comparative analysis is thorough and supported by citations.
- [ ] Mechanistic insights are clearly articulated.
- [ ] Limitations and alternative interpretations are transparently addressed.
- [ ] Clinical and broader implications are cohesively integrated.
--->

# Conclusion

-   Summarize the main conclusions of your study clearly and concisely.
-   Emphasize the importance and significance of your findings.
-   Avoid introducing new information or repeating details already discussed in the Results or Discussion sections.

# References

<div id="refs"></div>

# Author Contributions

Conceptualization, [...]; Data curation, [...]; Investigation, [...]; Methodology, [...]; Supervision, [...]; Formal analysis, [...]; Visualization, [...]; Writing–original draft, [...]; Writing–review & editing, [...]. All authors have read and agreed to the published version of the manuscript.

# Funding

This work was funded by [Funding institution] through [Project name] [Funding code/ID].

# Institutional Review Board Statement

Ethical approval was obtained from the Ethics Committee of the [Committee name] ([Document number]).

# Informed Consent Statement

All participants received detailed information regarding the study objectives, procedures, and potential implications. Informed consent was obtained to ensure ethical compliance and participant autonomy.

# Data availability statement

The raw data supporting the conclusions of this article will be made available by the authors without undue reservation.

# Conflicts of interests

The authors declare that the research was conducted without any commercial or financial relationships construed as as a potential conflict of interest.